Rapid-response Bayesian sky localization for electromagnetic follow-up of gravitational-wave candidates

### 2013-02-13 ms.tex
Added SNR signal model

In the frequency domain, the strain $\tilde x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
%
\begin{equation}\label{eq:signal-model}
	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
\end{equation}
%
where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base. The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be 1~Mpc,
%
\begin{equation}\label{eq:sigma}
    \sigma^2 = \int_0^\infty \frac{\left| \tilde u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
\end{equation}
%
The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
%
\begin{equation}\label{eq:gaussian-likelihood}
	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
		\propto \exp \left[
		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
	\right].
\end{equation}

@@

If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, \iota, \psi, \phi_c) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, inclination $\mathrm{\iota}$, polarization angle $\psi$, and coalescence phase $\phi_c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c) \right)^2
    \right]
\end{equation}

>>>>

In the frequency domain, the strain $\tilde x (\omega)$ observed by a detector is a linear combination of two orthogonal \ac{GW} quadratures and additive noise,
%
\begin{equation}\label{eq:signal-model}
	\tilde x (\omega) = \tilde h (\omega, \rho, \gamma, \tau) + \tilde n (\omega) = \frac{\rho}{\sigma} e^{\imath (\gamma - \omega \tau)} \tilde u (\omega) + \tilde n (\omega),
\end{equation}
%
where $\tilde h (\omega)$ is the signal content, $\tilde n (\omega)$ is the detector's \ac{WSS} Gaussian noise with \ac{PSD} $S(\omega)$, $\tilde u (\omega)$ is the frequency-domain post-Newtonian inspiral signal, $\rho$ is the signal's amplitude, $\gamma$ is a phase angle the encodes the relative strength of two \ac{GW} quadratures, and $\tau$ is the time of arrival relative to an arbitrary time base. The quantity $\sigma$ is a normalization constant defined as the \ac{SNR} of the \ac{GW} source at a fiducial distance, here taken to be $D_\mathrm{fiducial} = 1$~Mpc,
%
\begin{equation}\label{eq:sigma}
    \sigma^2 = \int_0^\infty \frac{\left| \tilde u (\omega)\right|^2}{S(\omega)} \, \mathrm{d}\omega.
\end{equation}
%
The distance at which an optimally oriented \ac{GW} source would produce \ac{SNR}=1 is defined as $D_1 = D_\mathrm{fiducial} / \sigma$. The likelihood, or the probability of obtaining the observation $\tilde x(\omega)$ conditioned on the value of the parameter vector $\boldsymbol\theta = (\rho, \gamma, \tau)$, is Gaussian, with the log-likelihood proportional to:
%
\begin{equation}\label{eq:gaussian-likelihood}
	\mathcal{L}(\tilde x; \boldsymbol\theta) = p(\tilde x | \boldsymbol\theta)
		\propto \exp \left[
		- \frac{1}{2} \int_0^\infty \frac{\left|\tilde x (\omega)
			- \tilde h (\omega; \boldsymbol\theta) \right|^2}{S(\omega)} \, \mathrm{d}\omega
	\right].
\end{equation}

@@

If we assume that the time and amplitude estimation errors are uncorrelated, then we can separate the likelihood into one factor that depends only on time and one that depends only on amplitude:
%
\begin{eqnarray*}
    \mathcal{L}(
        \hat{\tau}_1, \hat{\tau}_2, &\dots, \hat{\tau}_N,
        \hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
        \mathbf{n}, t_\oplus, D_\mathrm{L}, \iota, \psi, \phi_c) \\
    &= \mathcal{L}_\mathrm{TOA}(
        \hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N; \mathbf{n}, t_\oplus) \\
        &\times
        \mathcal{L}_\mathrm{SNR}(\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N;
         \mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c
    ).
\end{eqnarray*}
%
The \ac{TOA}-only factor depends only on the \ac{TOA} estimates $\hat{\tau}_1, \hat{\tau}_2, \dots, \hat{\tau}_N$, the unknown sky location $\mathbf{n}$ (or equivalently, the right ascension $\alpha$, declination $\delta$, and sidereal time), and the absolute \ac{TOA} at an arbitrary reference point (herein, the geocenter), $t_\oplus$, the fixed geographical locations of the detectors, $\mathbf{r}_i$, and the speed of light $c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{TOA} \propto
    \exp \left[
            -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i - \tau_i\right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
        \right]
    = \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \frac{\left(\hat{\tau}_i + \mathbf{n} \cdot \mathbf{r}_i / c - t_\oplus \right)^2}{\left(\hat{\rho_i} / \omega_{\mathrm{rms},i}\right)^2}
    \right]
\end{equation}
%
The \ac{SNR}-only factor depends only on the \ac{SNR} estimates $\hat{\rho}_1, \hat{\rho}_2, \dots, \hat{\rho}_N$, but it depends on the unknown sky location, luminosity distance $D_\mathrm{L}$, inclination $\mathrm{\iota}$, polarization angle $\psi$, and coalescence phase $\phi_c$.
%
\begin{equation}
    \mathcal{L}_\mathrm{SNR} \propto \exp \left[
        -\frac{1}{2} \sum_{i=1}^N \left(\hat{\rho}_i - \rho_i(\mathbf{n}, D_\mathrm{L}, \iota, \psi, \phi_c) \right)^2
    \right]
\end{equation}
%
It is easily shown that
%
\begin{eqnarray*}
    {\rho_i}^2 &=
        \frac{1}{8} \left(\frac{D_{1,i}}{D_\mathrm{L}}\right)^2 \left[
            ( {F_{+,i}}^2 + {F_{\times,i}}^2 \right)
            \left(1 + 6 x^2 + x^4 \right) \\
            &+ \left(1 - x^2 \right) \left( {F_{+,i}}^2 - {F_{\times,i}}^2 \right) \cos {2\psi}
            + 2 \left( {F_{+,i}} {F_{\times,i}} ) \sin {2\psi}
        \right],
\end{eqnarray*}
%
where $F_{+,i}$ and $F_{\times,i}$ are the `$+$' and `$\times$' antenna factors of detector $i$, themselves functions of sky location $\mathbf{n}$.
